### 2024-04-14 paper/paper.qmd
added new data plot

---
title: "Decoding Sleep Disorders Through Self-Reported Patterns"
subtitle: "A Logistic Regression Approach to the NHANES 2017-March 2020 Sleep Data"
title-block-banner: true
abstract: "This study investigates the potential relationship between individuals' sleep patterns and the self-reported incidence of sleep troubles as confirmed by a medical professional. Using logistic regression analysis on data from the NHANES 2017-March 2020 Sleep Disorders dataset, we examine how various factors such as sleep duration on weekdays and weekends, frequency of snoring, and daytime sleepiness are associated with the likelihood of reporting sleep problems to a doctor. The findings aim to shed light on the predictive value of self-monitored sleep behaviors in identifying individuals who may require medical attention for sleep-related issues. This analysis contributes to the broader understanding of sleep health and its complex interactions with daily functioning."
thanks: "Code and data supporting this analysis are available at https://github.com/TEJMaster/Sleep-Disorder-Analysis"
author: "Terry Tu"
date: today
date-format: long
format: pdf
toc: true
toc-depth: 1
number-sections: true
bibliography: references.bib
link-citations: true
editor: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE, message=FALSE}
## Import necessary libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
library(readr)
library(knitr)
library(kableExtra)
library(rstanarm)
library(haven)
library(arrow)
library(gridExtra)
```

# Introduction

The rhythm of our nightly rest is more than a personal habit; it's a public health pulse that captures the essence of well-being in our fast-paced society. The increasing prevalence of sleep disorders and their impact on daily life and overall health has become a point of societal concern, akin to the growing conversation around mental health and lifestyle diseases. The intimate link between quality of sleep and the vitality of an individual's life prompts a closer examination of sleep patterns within the populace. Drawing on the rich dataset from the National Health and Nutrition Examination Survey (NHANES) for the years 2017 to March 2020, our research delves into the self-reported instances of sleep disturbances and their association with various sleep behaviors.

This study is an explorative journey into the silent epidemic of sleep disorders that plague modern society, affecting productivity, mental health, and long-term wellbeing. The aim is to uncover the underlying patterns of sleep behavior that correlate with the reports of sleep troubles to medical professionals, thereby piecing together the nocturnal puzzle of restless societies. We explore the quantitative relationship between self-reported snoring frequency, feelings of daytime sleepiness, and the regularity of sleep hours during weekdays and weekends with the likelihood of reporting sleep issues to a doctor [@Prevention_2021].

Our analysis hinges on the application of logistic regression to the NHANES dataset, which presents a comprehensive view of American sleep habits. By interpreting the nuances of this rich dataset, we aspire to illuminate the factors that signal the need for medical attention in the domain of sleep health. The outcome of our investigation is poised to provide a scaffold for healthcare professionals and policymakers to base early intervention strategies, aiming to cultivate a well-rested population.

Following this introduction, the structure of the paper is laid out to facilitate a coherent flow of information and analysis. @sec-data (Data) provides a meticulous breakdown of the NHANES dataset, elucidating the data cleaning process and offering a descriptive overview of the key variables. @sec-model (Model) details the logistic regression model's design and the rationale behind the choice of predictors. @sec-result (Result) presents the findings, interpreted with precision and caution, alongside graphical representations for clarity. Concluding the paper, @sec-discussion (Discussion) reflects on the broader implications of the study, acknowledging limitations and proposing avenues for future research.

# Data {#sec-data}

```{r include=FALSE, message=FALSE, echo=FALSE}
# read the data file and store it in the 'data' variable
cleaned_data <- read_parquet(here("data", "analysis_data", "cleaned_data.parquet"))
```

## Raw Data

The dataset underpinning this analysis is derived from the National Health and Nutrition Examination Survey (NHANES), spanning from 2017 to March 2020. This public dataset includes responses from participants regarding their sleep patterns, incorporating 10,195 records initially. After meticulous data cleaning, the dataset for analysis stands at 10,031 records, encapsulating variables critical to our research: the respondent's ID, usual sleep and wake times on both weekdays and weekends, total sleep duration, frequency of snoring, incidence of breathing pauses during sleep, and self-reported communication of sleep troubles to a health professional. The dataset provides a snapshot of Americans' sleep behaviors before the disruption caused by the COVID-19 pandemic [@Prevention_2021].

The survey participants' ages range widely, reflecting the diversity of the American population. Variables are finely tuned to capture the multifaceted nature of sleep, encompassing aspects such as duration, disruptions, and subjective experiences of daytime sleepiness. The NHANES protocol ensures that this dataset is a robust and reliable source of information, adhering to stringent ethical standards and data collection methods, as detailed in the NHANES Analytic Guidelines. For further information on the data cleaning specifics and validation checks, please see the supplementary material in @sec-add-data.

## Data Analysis Tools

Our statistical exploration was conducted within the R programming environment [@r], leveraging its comprehensive ecosystem for data analysis. We utilized the tidyverse collection of R packages [@rTidyverse] to streamline our data processing tasks. The ggplot2 package [@rGgplot2] was instrumental in crafting insightful visualizations that articulated the intricate relationships within our data. The dplyr package [@rDplyr] provided a syntax that facilitated the manipulation and transformation of our dataset, enabling us to prepare the data effectively for logistic regression analysis. Data importation was efficiently handled by the readr package [@rReadr], known for its quick and user-friendly approach to reading tabular data. Navigational simplicity within our project's directories was achieved with the here package [@R-here], which reliably managed file paths without the need for manual path setting. The reproducibility of our research was enhanced by the knitr package [@rKnitr], which seamlessly wove R code into our report, ensuring that our findings are transparent and replicable. For tabular data presentation, kableExtra[@R-kableExtra] offered a suite of customization options that enhanced the readability and aesthetic appeal of our tables. The logistic regression model was developed using core functions in R, which provide robust methods for estimating the effects of various predictors on a binary outcome.

## Variable Description {#sec-variable}

**Weekday Sleep Duration (SLD012)**: This variable measures the total number of hours respondents usually sleep on weekdays or workdays, with values ranging from 2 to 14 hours. It provides insight into their sleep patterns during the typical workweek.

**Weekend Sleep Duration (SLD013)**: Similar to the weekday sleep duration, this variable represents the total number of hours respondents usually sleep on weekends or non-workdays, also ranging from 2 to 14 hours. It helps in understanding the variation in sleep patterns during days off from work.

**Snoring Frequency (SLQ030)**: This variable records how often respondents snore while sleeping, with responses ranging from 0 (Never) to 3 (Frequently). Snoring is a common symptom of sleep disorders such as obstructive sleep apnea, making this variable relevant to the study of sleep health.

**Overly Sleep Frequency (SLQ120)**: This variable assesses how often respondents feel excessively or overly sleepy during the day, with values ranging from 0 (Never) to 4 (Almost always). It is an indicator of sleep quality and quantity, as well as potential sleep disorders.

**Reported Sleep Trouble (SLQ050)**: The dependent variable in this study, it indicates whether respondents have ever told a doctor or other health professional that they have trouble sleeping. It is treated as a binary outcome variable, with values of 0 (No report of sleep trouble) and 1 (Reporting sleep trouble).

## Sample of Cleaned Sleep Disorder Data

```{r}
#| label: tbl-sleep-data-sample
#| tbl-cap: Sample of Sleep Disorder Data
#| echo: false
# Code created with the help from ChatGPT
sample_data <- head(cleaned_data) %>%
  select(RespondentID, WeekdaySleepDuration, WeekendSleepDuration, SnoringFrequency, BreathingPauseFrequency, OverlySleepFrequency, ReportedSleepTrouble)

# Create a nicely formatted table with custom column names
kable(sample_data, col.names = c("Respondent ID", "Weekday Sleep Duration (hrs)", "Weekend Sleep Duration (hrs)", "Snoring Frequency", "Breathing Pause Frequency", "Overly Sleep Frequency", "Reported Sleep Trouble"))

```

@tbl-sleep-data-sample represents a subset of the broader NHANES sleep disorder dataset. Each row in the table corresponds to an individual participant, uniquely identified by their Respondent ID. The "Weekday Sleep Duration (hrs)" and "Weekend Sleep Duration (hrs)" columns quantify the number of hours slept during the weekdays and weekends, respectively, providing a snapshot of the individual's sleep patterns. "Snoring Frequency" and "Breathing Pause Frequency" are categorical measures that reflect how often the respondents experience snoring and breathing pauses during sleep, common indicators of sleep disturbances such as sleep apnea. The "Overly Sleep Frequency" column indicates the frequency at which respondents report feeling overly sleepy during the day, a sign that can be indicative of inadequate sleep quality or quantity. Lastly, the "Reported Sleep Trouble" column is a binary measure showing whether the respondent has reported having sleep troubles to a health professional, with 0 signifying no reported trouble and 1 indicating reported trouble.

## Measurement:

In this study, we utilized data from the National Health and Nutrition Examination Survey (NHANES), specifically focusing on the sleep disorders component which includes data collected between 2017 and March 2020. The NHANES program, a longstanding project conducted by the National Center for Health Statistics (NCHS), plays a critical role in assessing the health and nutritional status of adults and children in the United States. This dataset is pivotal in understanding public health and informs policy decisions through scientifically reliable data [@Prevention_2021].

The sleep disorders dataset within NHANES is enriched by questions adapted from the Munich ChronoType Questionnaire [@Roenneberg2003], targeting various aspects of sleep behavior and disorders. The inclusion of these questions is instrumental in exploring the complex dynamics of sleep patterns among the U.S. population. Due to disruptions caused by the COVID-19 pandemic, the 2019-2020 data collection cycle was prematurely halted in March 2020, leading to its combination with the 2017-2018 cycle to ensure national representativeness and analytical robustness.

This combined dataset referred to as the NHANES 2017-March 2020 pre-pandemic data, offers valuable insights into the sleep habits of Americans before the pandemic. It is instrumental for researchers and public health officials aiming to understand baseline sleep behaviors and potential disturbances across a broad demographic spectrum.

To prepare the dataset for analysis, extensive data cleaning and processing were conducted. This included removing entries with missing, refused, or 'don't know' responses for critical variables such as sleep duration and frequency of snoring. Additionally, to address issues with data reliability and consistency, about 3% of the data underwent verification through audio recordings of the interviews. Moreover, for variables capturing sleep duration on weekdays (SLD012) and weekends (SLD013), reported times were meticulously reviewed, with outliers adjusted and rounded to the nearest half-hour, enhancing the data's accuracy and usability.

Detailed descriptions of the variables used in this study, along with the specific adjustments made to the dataset, are available in  @sec-variable. This section is designed to provide a comprehensive understanding of the origins, processing, and analytical framework applied to each variable relevant to this study.

## Data Exploration: {#sec-exploration}

In this section, we explore the distributions of key variables related to sleep patterns and disorders in the NHANES dataset. Histograms provide visual insights into the frequency of reported sleep duration, snoring, breathing pauses, daytime sleepiness, and reported sleep troubles.

```{r fig-histograms, echo=FALSE, fig.width=8, fig.height=6, fig.cap="Distributions of sleep-related variables"}
# Generate individual histograms for each variable
p1 <- ggplot(cleaned_data, aes(x = WeekdaySleepDuration)) + 
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") + 
  labs(x = "Weekday Sleep Duration (hrs)", y = "Count") +
  theme_minimal()

p2 <- ggplot(cleaned_data, aes(x = WeekendSleepDuration)) + 
  geom_histogram(binwidth = 1, fill = "salmon", color = "black") + 
  labs(x = "Weekend Sleep Duration (hrs)", y = "Count") +
  theme_minimal()

p3 <- ggplot(cleaned_data, aes(x = SnoringFrequency)) + 
  geom_histogram(binwidth = 1, fill = "lightgreen", color = "black") + 
  labs(x = "Snoring Frequency", y = "Count") +
  theme_minimal()

p4 <- ggplot(cleaned_data, aes(x = BreathingPauseFrequency)) + 
  geom_histogram(binwidth = 1, fill = "plum", color = "black") + 
  labs(x = "Breathing Pause Frequency", y = "Count") +
  theme_minimal()

p5 <- ggplot(cleaned_data, aes(x = OverlySleepFrequency)) + 
  geom_histogram(binwidth = 1, fill = "gold", color = "black") + 
  labs(x = "Overly Sleep Frequency", y = "Count") +
  theme_minimal()

p6 <- ggplot(cleaned_data, aes(x = as.factor(ReportedSleepTrouble))) + 
  geom_bar(fill = "lightblue", color = "black") + 
  labs(x = "Reported Sleep Trouble", y = "Count") +
  scale_x_discrete(labels = c("0" = "No Trouble", "1" = "Trouble")) +
  theme_minimal()

# Arrange the plots into a 2x3 grid
grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 3)
```

**Weekday Sleep Duration (hrs)**: The histogram for weekday sleep duration exhibits a unimodal distribution, peaking around 7 to 8 hours, which aligns with the generally recommended sleep duration for adults. The symmetrical shape suggests a commonality in sleep patterns among individuals during the workweek.

**Weekend Sleep Duration (hrs)**: Similarly unimodal, the weekend sleep duration distribution also peaks around the same range as weekday sleep but shows a noticeable tendency for slightly longer durations. This may suggest that individuals take the opportunity to sleep more on weekends, possibly compensating for the workweek.

**Snoring Frequency**: The snoring frequency histogram reveals a broad distribution, with the majority of respondents indicating they never or rarely snore. A significant count of individuals also reports frequent snoring, suggesting that the experience of snoring is quite varied among the respondents.

**Breathing Pause Frequency**: The distribution for breathing pause frequency is steeply skewed towards 'Never', suggesting that breathing pauses during sleep are infrequently experienced or reported by the majority of respondents. The scant instances of frequent breathing pauses may indicate a lower occurrence or a lack of self-awareness of these events.

**Overly Sleep Frequency**: The distribution of daytime sleepiness frequency is right-skewed, with 'Sometimes (2-4 times a month)' emerging as the modal category. However, a notable proportion of respondents experience sleepiness during the day at least once a month, suggesting that daytime sleepiness is a common issue.

**Reported Sleep Trouble**: The binary distribution for reported sleep trouble shows a larger proportion of individuals reporting no sleep trouble compared to those reporting trouble, approximately threefold. This ratio highlights that while sleep trouble is present, it is not reported by the majority of participants in this sample.

# Model {#sec-model}

The aim of our model is to explore the relationship between various sleep-related factors and the self-reported incidence of sleep troubles. We employ a Bayesian logistic regression model to analyze the data from the National Health and Nutrition Examination Survey (NHANES). Further details and diagnostics of this model are provided in @sec-posterior.

## Model set-up

Let $y_i$ denote the binary outcome indicating whether an individual has reported sleep troubles to a doctor, with $y_i = 1$ for reported troubles and $y_i = 0$ otherwise. The predictors include weekday sleep duration ($x_{i1}$), weekend sleep duration ($x_{i2}$), snoring frequency ($x_{i3}$), breathing pause frequency ($x_{i4}$), and daytime sleepiness ($x_{i5}$). The logistic regression model is formulated as follows:

\begin{align}
y_i | p_i &\sim \mbox{Bernoulli}(p_i) \\
\mbox{logit}(p_i) &= \beta_0 + \beta_1 \times x_{i1} + \beta_2 \times x_{i2} + \beta_3 \times x_{i3} + \beta_4 \times x_{i4} + \beta_5 \times x_{i5} \\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1, \beta_2, \beta_3, \beta_4, \beta_5 &\sim \mbox{Normal}(0, 2.5)
\end{align}

The model is implemented in R using the rstanarm package, enabling the incorporation of prior beliefs about the parameters and estimation of their posterior distributions. The stan_glm function from rstanarm is used to perform the Bayesian logistic regression, integrating both the data and prior knowledge for inference.

Priors for the intercept and coefficients are chosen to reflect a neutral stance before observing the data, centered around zero with a moderate spread to allow flexibility in the estimates. These priors convey an expectation of no substantial effect from each predictor, with the data informing deviations from this baseline assumption.

The logistic regression framework models the log odds of reporting sleep troubles as a linear combination of the predictors, with the coefficients indicating the direction and magnitude of each predictor's effect. The stan_glm function combines these priors with the data to estimate the posterior distributions of the model parameters, providing a comprehensive view of the relationships between sleep-related factors and reported sleep troubles.

## Model Justification
In our analysis, we hypothesize that certain sleep-related factors will have significant associations with the likelihood of reporting sleep troubles. Specifically:

**Weekday Sleep Duration ($\beta_1$)**: We hypothesize that this coefficient may be positive, indicating that longer sleep durations on weekdays could be associated with a higher likelihood of reporting sleep troubles. This could reflect the possibility that individuals with sleep disturbances may attempt to compensate by sleeping longer.

**Weekend Sleep Duration ($\beta_2$)**: Conversely, we hypothesize that this coefficient may be negative, suggesting that longer sleep durations on weekends might be associated with a lower likelihood of reporting sleep troubles. This could indicate that catching up on sleep during weekends may alleviate some sleep-related issues.

**Snoring Frequency ($\beta_3$)**: We expect this coefficient to be positive, as frequent snoring is often a symptom of sleep disorders such as obstructive sleep apnea, which could lead to reporting sleep troubles.

**Breathing Pause Frequency ($\beta_4$)**: Similarly, we hypothesize that this coefficient will be positive, as pauses in breathing during sleep are a hallmark of sleep apnea, a condition commonly associated with sleep disturbances.

**Daytime Sleepiness ($\beta_5$)**: We anticipate that this coefficient will also be positive, as excessive sleepiness during the day is a common consequence of poor sleep quality or insufficient sleep, which may prompt individuals to report sleep troubles.

The Bayesian framework allows us to incorporate these hypotheses as priors in our model, providing a structured way to integrate our substantive expectations with the data analysis. By doing so, we can obtain a more informed and nuanced interpretation of the relationships between sleep-related factors and the likelihood of reporting sleep troubles.

# Result {#sec-result}

# Discussion {#sec-discussion}

# Appendix {#sec-appendix}

## Additional Data Details {#sec-add-data}

## Model Details

### Posterior Predictive Check {#sec-posterior}

# References