### 2024-04-14 paper/paper.qmd
added result

---
title: "Decoding Sleep Disorders Through Self-Reported Patterns"
subtitle: "A Logistic Regression Approach to the NHANES 2017-March 2020 Sleep Data"
title-block-banner: true
abstract: "This study investigates the potential relationship between individuals' sleep patterns and the self-reported incidence of sleep troubles as confirmed by a medical professional. Using logistic regression analysis on data from the NHANES 2017-March 2020 Sleep Disorders dataset, we examine how various factors such as sleep duration on weekdays and weekends, frequency of snoring, and daytime sleepiness are associated with the likelihood of reporting sleep problems to a doctor. The findings aim to shed light on the predictive value of self-monitored sleep behaviors in identifying individuals who may require medical attention for sleep-related issues. This analysis contributes to the broader understanding of sleep health and its complex interactions with daily functioning."
thanks: "Code and data supporting this analysis are available at https://github.com/TEJMaster/Sleep-Disorder-Analysis"
author: "Terry Tu"
date: today
date-format: long
format: pdf
toc: true
toc-depth: 1
number-sections: true
bibliography: references.bib
link-citations: true
editor: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE, message=FALSE}
## Import necessary libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
library(readr)
library(knitr)
library(kableExtra)
library(rstanarm)
library(haven)
library(arrow)
library(gridExtra)
library(modelsummary)
```

# Introduction

The rhythm of our nightly rest is more than a personal habit; it's a public health pulse that captures the essence of well-being in our fast-paced society. The increasing prevalence of sleep disorders and their impact on daily life and overall health has become a point of societal concern, akin to the growing conversation around mental health and lifestyle diseases. The intimate link between quality of sleep and the vitality of an individual's life prompts a closer examination of sleep patterns within the populace. Drawing on the rich dataset from the National Health and Nutrition Examination Survey (NHANES) for the years 2017 to March 2020, our research delves into the self-reported instances of sleep disturbances and their association with various sleep behaviors.

This study is an explorative journey into the silent epidemic of sleep disorders that plague modern society, affecting productivity, mental health, and long-term wellbeing. The aim is to uncover the underlying patterns of sleep behavior that correlate with the reports of sleep troubles to medical professionals, thereby piecing together the nocturnal puzzle of restless societies. We explore the quantitative relationship between self-reported snoring frequency, feelings of daytime sleepiness, and the regularity of sleep hours during weekdays and weekends with the likelihood of reporting sleep issues to a doctor [@Prevention_2021].

Our analysis hinges on the application of logistic regression to the NHANES dataset, which presents a comprehensive view of American sleep habits. By interpreting the nuances of this rich dataset, we aspire to illuminate the factors that signal the need for medical attention in the domain of sleep health. The outcome of our investigation is poised to provide a scaffold for healthcare professionals and policymakers to base early intervention strategies, aiming to cultivate a well-rested population.

Following this introduction, the structure of the paper is laid out to facilitate a coherent flow of information and analysis. @sec-data (Data) provides a meticulous breakdown of the NHANES dataset, elucidating the data cleaning process and offering a descriptive overview of the key variables. @sec-model (Model) details the logistic regression model's design and the rationale behind the choice of predictors. @sec-result (Result) presents the findings, interpreted with precision and caution, alongside graphical representations for clarity. Concluding the paper, @sec-discussion (Discussion) reflects on the broader implications of the study, acknowledging limitations and proposing avenues for future research.

# Data {#sec-data}

```{r include=FALSE, message=FALSE, echo=FALSE}
# read the data file and store it in the 'data' variable
cleaned_data <- read_parquet(here("data", "analysis_data", "cleaned_data.parquet"))
```

## Raw Data

The dataset underpinning this analysis is derived from the National Health and Nutrition Examination Survey (NHANES), spanning from 2017 to March 2020. This public dataset includes responses from participants regarding their sleep patterns, incorporating 10,195 records initially. After meticulous data cleaning, the dataset for analysis stands at 10,031 records, encapsulating variables critical to our research: the respondent's ID, usual sleep and wake times on both weekdays and weekends, total sleep duration, frequency of snoring, incidence of breathing pauses during sleep, and self-reported communication of sleep troubles to a health professional. The dataset provides a snapshot of Americans' sleep behaviors before the disruption caused by the COVID-19 pandemic [@Prevention_2021].

The survey participants' ages range widely, reflecting the diversity of the American population. Variables are finely tuned to capture the multifaceted nature of sleep, encompassing aspects such as duration, disruptions, and subjective experiences of daytime sleepiness. The NHANES protocol ensures that this dataset is a robust and reliable source of information, adhering to stringent ethical standards and data collection methods, as detailed in the NHANES Analytic Guidelines. For further information on the data cleaning specifics and validation checks, please see the supplementary material in @sec-add-data.

## Data Analysis Tools

Our statistical exploration was conducted within the R programming environment [@r], leveraging its comprehensive ecosystem for data analysis. We utilized the tidyverse collection of R packages [@rTidyverse] to streamline our data processing tasks. The ggplot2 package [@rGgplot2] was instrumental in crafting insightful visualizations that articulated the intricate relationships within our data. The dplyr package [@rDplyr] provided a syntax that facilitated the manipulation and transformation of our dataset, enabling us to prepare the data effectively for logistic regression analysis. Data importation was efficiently handled by the readr package [@rReadr], known for its quick and user-friendly approach to reading tabular data. Navigational simplicity within our project's directories was achieved with the here package [@R-here], which reliably managed file paths without the need for manual path setting. The reproducibility of our research was enhanced by the knitr package [@rKnitr], which seamlessly wove R code into our report, ensuring that our findings are transparent and replicable. For tabular data presentation, kableExtra[@R-kableExtra] offered a suite of customization options that enhanced the readability and aesthetic appeal of our tables. The logistic regression model was developed using core functions in R, which provide robust methods for estimating the effects of various predictors on a binary outcome.

## Variable Description {#sec-variable}

**Weekday Sleep Duration (SLD012)**: This variable measures the total number of hours respondents usually sleep on weekdays or workdays, with values ranging from 2 to 14 hours. It provides insight into their sleep patterns during the typical workweek.

**Weekend Sleep Duration (SLD013)**: Similar to the weekday sleep duration, this variable represents the total number of hours respondents usually sleep on weekends or non-workdays, also ranging from 2 to 14 hours. It helps in understanding the variation in sleep patterns during days off from work.

**Snoring Frequency (SLQ030)**: This variable records how often respondents snore while sleeping, with responses ranging from 0 (Never) to 3 (Frequently). Snoring is a common symptom of sleep disorders such as obstructive sleep apnea, making this variable relevant to the study of sleep health.

**Overly Sleep Frequency (SLQ120)**: This variable assesses how often respondents feel excessively or overly sleepy during the day, with values ranging from 0 (Never) to 4 (Almost always). It is an indicator of sleep quality and quantity, as well as potential sleep disorders.

**Reported Sleep Trouble (SLQ050)**: The dependent variable in this study, it indicates whether respondents have ever told a doctor or other health professional that they have trouble sleeping. It is treated as a binary outcome variable, with values of 0 (No report of sleep trouble) and 1 (Reporting sleep trouble).

## Sample of Cleaned Sleep Disorder Data

```{r}
#| label: tbl-sleep-data-sample
#| tbl-cap: Sample of Sleep Disorder Data
#| echo: false
# Code created with the help from ChatGPT
sample_data <- head(cleaned_data) %>%
  select(RespondentID, WeekdaySleepDuration, WeekendSleepDuration, SnoringFrequency, BreathingPauseFrequency, OverlySleepFrequency, ReportedSleepTrouble)

# Create a nicely formatted table with custom column names
kable(sample_data, col.names = c("Respondent ID", "Weekday Sleep Duration (hrs)", "Weekend Sleep Duration (hrs)", "Snoring Frequency", "Breathing Pause Frequency", "Overly Sleep Frequency", "Reported Sleep Trouble"))

```

@tbl-sleep-data-sample represents a subset of the broader NHANES sleep disorder dataset. Each row in the table corresponds to an individual participant, uniquely identified by their Respondent ID. The "Weekday Sleep Duration (hrs)" and "Weekend Sleep Duration (hrs)" columns quantify the number of hours slept during the weekdays and weekends, respectively, providing a snapshot of the individual's sleep patterns. "Snoring Frequency" and "Breathing Pause Frequency" are categorical measures that reflect how often the respondents experience snoring and breathing pauses during sleep, common indicators of sleep disturbances such as sleep apnea. The "Overly Sleep Frequency" column indicates the frequency at which respondents report feeling overly sleepy during the day, a sign that can be indicative of inadequate sleep quality or quantity. Lastly, the "Reported Sleep Trouble" column is a binary measure showing whether the respondent has reported having sleep troubles to a health professional, with 0 signifying no reported trouble and 1 indicating reported trouble.

## Measurement:

In this study, we utilized data from the National Health and Nutrition Examination Survey (NHANES), specifically focusing on the sleep disorders component which includes data collected between 2017 and March 2020. The NHANES program, a longstanding project conducted by the National Center for Health Statistics (NCHS), plays a critical role in assessing the health and nutritional status of adults and children in the United States. This dataset is pivotal in understanding public health and informs policy decisions through scientifically reliable data [@Prevention_2021].

The sleep disorders dataset within NHANES is enriched by questions adapted from the Munich ChronoType Questionnaire [@Roenneberg2003], targeting various aspects of sleep behavior and disorders. The inclusion of these questions is instrumental in exploring the complex dynamics of sleep patterns among the U.S. population. Due to disruptions caused by the COVID-19 pandemic, the 2019-2020 data collection cycle was prematurely halted in March 2020, leading to its combination with the 2017-2018 cycle to ensure national representativeness and analytical robustness.

This combined dataset referred to as the NHANES 2017-March 2020 pre-pandemic data, offers valuable insights into the sleep habits of Americans before the pandemic. It is instrumental for researchers and public health officials aiming to understand baseline sleep behaviors and potential disturbances across a broad demographic spectrum.

To prepare the dataset for analysis, extensive data cleaning and processing were conducted. This included removing entries with missing, refused, or 'don't know' responses for critical variables such as sleep duration and frequency of snoring. Additionally, to address issues with data reliability and consistency, about 3% of the data underwent verification through audio recordings of the interviews. Moreover, for variables capturing sleep duration on weekdays (SLD012) and weekends (SLD013), reported times were meticulously reviewed, with outliers adjusted and rounded to the nearest half-hour, enhancing the data's accuracy and usability.

Detailed descriptions of the variables used in this study, along with the specific adjustments made to the dataset, are available in  @sec-variable. This section is designed to provide a comprehensive understanding of the origins, processing, and analytical framework applied to each variable relevant to this study.

## Data Exploration: {#sec-exploration}

In this section, we explore the distributions of key variables related to sleep patterns and disorders in the NHANES dataset. Histograms provide visual insights into the frequency of reported sleep duration, snoring, breathing pauses, daytime sleepiness, and reported sleep troubles.

```{r fig-histograms, echo=FALSE, fig.width=8, fig.height=6, fig.cap="Distributions of sleep-related variables"}
# Generate individual histograms for each variable
p1 <- ggplot(cleaned_data, aes(x = WeekdaySleepDuration)) + 
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") + 
  labs(x = "Weekday Sleep Duration (hrs)", y = "Count") +
  theme_minimal()

p2 <- ggplot(cleaned_data, aes(x = WeekendSleepDuration)) + 
  geom_histogram(binwidth = 1, fill = "salmon", color = "black") + 
  labs(x = "Weekend Sleep Duration (hrs)", y = "Count") +
  theme_minimal()

p3 <- ggplot(cleaned_data, aes(x = SnoringFrequency)) + 
  geom_histogram(binwidth = 1, fill = "lightgreen", color = "black") + 
  labs(x = "Snoring Frequency", y = "Count") +
  theme_minimal()

p4 <- ggplot(cleaned_data, aes(x = BreathingPauseFrequency)) + 
  geom_histogram(binwidth = 1, fill = "plum", color = "black") + 
  labs(x = "Breathing Pause Frequency", y = "Count") +
  theme_minimal()

p5 <- ggplot(cleaned_data, aes(x = OverlySleepFrequency)) + 
  geom_histogram(binwidth = 1, fill = "gold", color = "black") + 
  labs(x = "Overly Sleep Frequency", y = "Count") +
  theme_minimal()

p6 <- ggplot(cleaned_data, aes(x = as.factor(ReportedSleepTrouble))) + 
  geom_bar(fill = "lightblue", color = "black") + 
  labs(x = "Reported Sleep Trouble", y = "Count") +
  scale_x_discrete(labels = c("0" = "No Trouble", "1" = "Trouble")) +
  theme_minimal()

# Arrange the plots into a 2x3 grid
grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 3)
```

**Weekday Sleep Duration (hrs)**: The histogram for weekday sleep duration exhibits a unimodal distribution, peaking around 7 to 8 hours, which aligns with the generally recommended sleep duration for adults. The symmetrical shape suggests a commonality in sleep patterns among individuals during the workweek.

**Weekend Sleep Duration (hrs)**: Similarly unimodal, the weekend sleep duration distribution also peaks around the same range as weekday sleep but shows a noticeable tendency for slightly longer durations. This may suggest that individuals take the opportunity to sleep more on weekends, possibly compensating for the workweek.

**Snoring Frequency**: The snoring frequency histogram reveals a broad distribution, with the majority of respondents indicating they never or rarely snore. A significant count of individuals also reports frequent snoring, suggesting that the experience of snoring is quite varied among the respondents.

**Breathing Pause Frequency**: The distribution for breathing pause frequency is steeply skewed towards 'Never', suggesting that breathing pauses during sleep are infrequently experienced or reported by the majority of respondents. The scant instances of frequent breathing pauses may indicate a lower occurrence or a lack of self-awareness of these events.

**Overly Sleep Frequency**: The distribution of daytime sleepiness frequency is right-skewed, with 'Sometimes (2-4 times a month)' emerging as the modal category. However, a notable proportion of respondents experience sleepiness during the day at least once a month, suggesting that daytime sleepiness is a common issue.

**Reported Sleep Trouble**: The binary distribution for reported sleep trouble shows a larger proportion of individuals reporting no sleep trouble compared to those reporting trouble, approximately threefold. This ratio highlights that while sleep trouble is present, it is not reported by the majority of participants in this sample.

# Model {#sec-model}

The aim of our model is to explore the relationship between various sleep-related factors and the self-reported incidence of sleep troubles. We employ a Bayesian logistic regression model to analyze the data from the National Health and Nutrition Examination Survey (NHANES). Further details and diagnostics of this model are provided in @sec-posterior.

## Model set-up

Let $y_i$ denote the binary outcome indicating whether an individual has reported sleep troubles to a doctor, with $y_i = 1$ for reported troubles and $y_i = 0$ otherwise. The predictors include weekday sleep duration ($x_{i1}$), weekend sleep duration ($x_{i2}$), snoring frequency ($x_{i3}$), breathing pause frequency ($x_{i4}$), and daytime sleepiness ($x_{i5}$). The logistic regression model is formulated as follows:

\begin{align}
y_i | p_i &\sim \mbox{Bernoulli}(p_i) \\
\mbox{logit}(p_i) &= \beta_0 + \beta_1 \times x_{i1} + \beta_2 \times x_{i2} + \beta_3 \times x_{i3} + \beta_4 \times x_{i4} + \beta_5 \times x_{i5} \\
\beta_0 &\sim \mbox{Normal}(0, 1) \\ % Assuming the log-odds of the base rate of reported sleep trouble is centered around 0
\beta_1, \beta_2, \beta_3, \beta_4, \beta_5 &\sim \mbox{Normal}(0, 1) \\ % Weekday sleep duration effect and Weekend sleep duration effect, Snoring frequency,  Breathing pause frequency, and Overly sleep frequency
\end{align}

The logistic regression model is defined using a Bayesian framework, implemented in R with the `rstanarm` package. This approach allows us to incorporate prior knowledge about the parameters and to estimate their posterior distributions based on the observed data.

In our model, the probability of an individual reporting sleep troubles, denoted by $p_i$, follows a Bernoulli distribution. The log odds of reporting sleep troubles are modeled as a linear combination of the predictors, with $\beta_0$ representing the intercept, and $\beta_1$ through $\beta_5$ representing the slopes for the respective predictors.

The priors for the intercept ($\beta_0$) and the slopes ($\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$) are deliberately chosen to reflect initial neutrality or skepticism regarding the impact of each predictor on the likelihood of reporting sleep troubles. Specifically, $\beta_0$ is assigned a Normal(0, 1) prior, indicating our initial uncertainty about the base rate of sleep trouble in the population.

For $\beta_1$ and $\beta_2$, which correspond to the effects of weekday and weekend sleep duration, we posit Normal(0, 0.5) priors. These priors are centered around zero, reflecting the number of sleep hour can be either positively or negatively contribute to sleep troubles, with a modest expectation of effect size.

The coefficients $\beta_3$, $\beta_4$, $\beta_5$ relating to snoring frequency, breathing pause frequency, and overly sleep frequency are each given a Normal(0, 1) prior. These priors imply that, prior to analyzing the data, we do not have strong expectations about the direction or scale of these effects.

## Model Justification

Our Bayesian logistic regression model is designed to investigate the associations between various sleep-related factors and the likelihood of reporting sleep troubles. The priors for the coefficients are set with specific hypotheses in mind, informed by existing literature and plausible biological mechanisms:

**Weekday Sleep Duration ($\beta_1$)**: We initially hypothesize that there may be a complex relationship between sleep duration on weekdays and sleep troubles. Long sleep durations could indicate an attempt to compensate for poor quality sleep, potentially related to sleep disturbances. As such, the prior for this coefficient is centered around zero, allowing the data to indicate the direction of association.

**Weekend Sleep Duration ($\beta_2$)**: For sleep duration on weekends, we consider the possibility of a recuperative effect, where catching up on sleep may reduce the likelihood of reporting troubles. This hypothesis is tentative, hence a prior centered around zero with a conservative standard deviation, permitting data-driven insights into this relationship.

**Snoring Frequency ($\beta_3$)**: Frequent snoring is a recognized symptom of sleep disorders like obstructive sleep apnea, which can lead to disrupted sleep. We posit a positive association between snoring frequency and reported sleep troubles, reflected in the priors for this coefficient.

**Breathing Pause Frequency ($\beta_4$)**: Breathing pauses during sleep are indicative of sleep apnea, which is directly related to sleep troubles. We hypothesize a positive association for this coefficient, which is incorporated into the prior setting.

**Daytime Sleepiness ($\beta_5$)**: Excessive daytime sleepiness is often the result of inadequate or disrupted nighttime sleep. We hypothesize a positive relationship between daytime sleepiness and the reporting of sleep troubles, expecting that those experiencing higher levels of sleepiness will be more likely to report troubles.

These hypotheses are structured into our Bayesian model as informed priors, which, while reflecting our initial expectations, are sufficiently flexible to be shaped by the empirical data. This integration of prior beliefs and observed information will yield posterior distributions that convey a robust interpretation of how sleep behaviors relate to the reporting of sleep troubles.

# Result {#sec-result}

```{r echo=FALSE}
# Retrieve the Bayesian logistic model from the saved .rds file
bayesian_logistic_model <- readRDS(file = here("models", "bayesian_logistic_model.rds"))
```

## Model Coefficients Interpretation

```{r}
#| label: tbl-logistic-coefficients
#| tbl-cap: "Summary Statistics for the Coefficients of the Logistic Model"
#| echo: false
# Round the coefficients directly when creating the data frame
coefficients_df <- data.frame(
  Term = names(coef(bayesian_logistic_model)),
  Estimate = round(coef(bayesian_logistic_model), 3)
)

# Create the kable without row names
coefficient_table <- kable(coefficients_df, 
                           col.names = c("Term", "Estimate"), 
                           align = 'c',
                           row.names = FALSE)  # Ensure row names are not printed

# Print the summary table
coefficient_table
```
@tbl-logistic-coefficients presents the key coefficients for understanding the influence of various sleep-related variables on the likelihood of reporting sleep troubles:

**Intercept ($\beta_0$)**:
The intercept of the model, $\beta_0$, is estimated at -1.396. This value represents the log odds of reporting sleep troubles when all the predictors are held at zero. It suggests that, in the absence of any sleep issues or without considering any sleep behaviors, the baseline log odds of reporting sleep troubles to a doctor are negative, indicating a lower probability of reporting sleep troubles.

**Weekday Sleep Duration ($\beta_1$)**:
For the weekday sleep duration, the coefficient $\beta_1$ is 0.051, suggesting a slight increase in the log odds of reporting sleep troubles for each additional hour of sleep during the weekdays. This could be indicative of a situation where those who sleep more on weekdays might be doing so due to sleep issues that have led them to consult with a doctor.

**Weekend Sleep Duration ($\beta_2$)**:
Conversely, the coefficient for weekend sleep duration, $\beta_2$, is -0.131. This implies that for each additional hour of sleep during the weekends, there is a decrease in the log odds of reporting sleep troubles. It could be interpreted as those who manage to sleep more on weekends are less likely to report sleep problems, possibly because catching up on sleep may alleviate some of their weekday sleep deficits.

**Snoring Frequency ($\beta_3$)**:
The coefficient for snoring frequency, $\beta_3$, has a value of 0.032. While this is a positive value, suggesting that an increase in snoring frequency is associated with an increase in the likelihood of reporting sleep troubles, the effect size is relatively small.

**Breathing Pause Frequency ($\beta_4$)**:
The coefficient for breathing pause frequency, $\beta_4$, is 0.412, indicating a more substantial positive association with reporting sleep troubles. This aligns with clinical understanding, as breathing pauses are often associated with sleep disorders like sleep apnea, which can be a significant concern prompting medical consultation.

**Overly Sleep Frequency ($\beta_5$)**:
Lastly, the coefficient for overly sleep frequency, $\beta_5$, is 0.425, suggesting a strong relationship with the reporting of sleep troubles. This result is intuitive, as feeling excessively sleepy can be a direct symptom of poor sleep quality or a sleep disorder, leading to discussions with a healthcare provider.

**Interpretation**:
The model indicates that various factors are related to the reporting of sleep troubles. While longer weekday sleep duration slightly increases the log odds of reporting sleep troubles, longer weekend sleep can reduce it. Frequent snoring has a smaller effect compared to breathing pauses and daytime sleepiness, which both have significant positive associations with reporting sleep troubles. These findings underline the complexity of sleep behavior and its impact on sleep quality and health consultations.

## Model Equation

\begin{align}
\text{logit}(p_i) &= -1.396 + 0.051 \times x_{i1} - 0.131 \times x_{i2} + 0.032 \times x_{i3} + 0.412 \times x_{i4} + 0.425 \times x_{i5} + \varepsilon_i
\end{align}

# Discussion {#sec-discussion}

# Appendix {#sec-appendix}

## Additional Data Details {#sec-add-data}

## Model Details

### Posterior Predictive Check {#sec-posterior}

# References